Modeling defoliation of Pinus Radiata trees using hyperspectral remote sensing data

### 2018-08-03 01_writing/Schratz_hyperspectral.tex
add xgboost 7 vars only to table and update results

The \textit{Gain} features is the most important one among the three. 
All measures sum up to one.
Downsides of this approach are that due to the contribution of three different parts it is complicated to understand why a specific feature was selected.
Furthermore the importance calculation approach is only valid for this algorithm and cannot be compared to others.
Nevertheless, as we only relied on the variable importance for this specific algorithm, using the internal xgboost approach was sufficient for this work.

@@

\begin{table}[t!]
\centering
\caption[t]{Predictive performance of \textit{xgboost} using all observations (All Observations) and observations from specific plots only (Single Plot Observations) with \ac{RMSE} as the error measure. The performance estimates for "All Observations" correspond to the fold for which the respective plot was serving as the test set. Column "Single Plot Observations", shows the mean performances at the repetition level of a spatial CV (5 folds, 5 repetitions), scored by using data of the respective plot only.}
\begingroup\footnotesize
\begin{tabular}{llll}
	\\
	Plot/Data & All Observations/all variables (Block CV) & All Observations/7 variables (Block CV) & Single Plot Observations (SpCV) \\
	\hline
	Laukiz 1  & 22.03                                     &                                         & 19.18                           \\
	Laukiz 2  & 51.75                                     &                                         & 17.24                           \\
	Luiando   & 13.20                                     &                                         & 8.30                            \\
	Oiartzun  & 32.97                                     &                                         & 14.40                           \\
	\bottomrule
\end{tabular}
\endgroup
\label{tab:supermodel_performance}
\end{table}

@@

@@

@@

@@

It is expected that models that were trained on the plot level only achieve a better performance than the global model.
The low performance on Laukiz 2 for the super model is most likely due to the difference of this plot to all others: The fitted model on Laukiz 1, Luiando and Oiartzun is not capable of reaching a good performance on the Laukiz 2 prediction data.
This is not surprising as Laukiz 2 shows substantially different plot characteristics compared to all others plots in terms of CV, mean point density (\autoref{arg}) and defoliation (\autoref{fig:defol_boxplots}).

@@

>>>>

The \textit{Gain} features is the most important one among the three.
All measures sum up to one (REFERENCE!).

@@

\begin{table}[t!]
\centering
\caption[t]{Predictive performance of \textit{xgboost} using all observations (All Observations) and observations from specific plots only (Single Plot Observations) with \ac{RMSE} as the error measure. The performance estimates for "All Observations" correspond to the fold for which the respective plot was serving as the test set. Column "Single Plot Observations", shows the mean performances at the repetition level of a spatial CV (5 folds, 5 repetitions), scored by using data of the respective plot only.}
\begingroup\footnotesize
\begin{tabular}{llll}
	\\
	Plot/Data & \specialcell{All Observations/ \\{all variables (Block CV)}} & \specialcell{All Observations/         \\ {7 variables (Block CV)}} & \specialcell{Plot level observations/\\{all variables (SpCV)}} \\
	\hline
	Laukiz 1  & 22.03                                                                              &       21.47                         & 19.18 \\
	Laukiz 2  & 51.75                                                                              &        49.94                        & 17.24 \\
	Luiando   & 13.20                                                                              &       15.37                        & 8.30  \\
	Oiartzun  & 32.97                                                                              &    17.62                            & 14.40 \\
	\bottomrule
\end{tabular}
\endgroup
\label{tab:supermodel_performance}
\end{table}

@@

Using only the seven most important variables (\autoref{fig:var-imp}) shows small increases in performance for Laukiz 1 and Laukiz 2, a small decrease for Luiando and almost a reduction of 50\% of the error for Oiartzun (32.97 vs. 17.62 RMSE).

@@

The best NRI scored rank eight (band 112 and band 62).
All further places up to rank 30 are NRIs.

@@

An important point that needs to be considered when interpreting the performance results is that we only related defoliation to indices derived from remote sensing data.
Possible other cofounding factors that could possibly help in predicting defoliation were not considered.
One example here is tree age: The older a tree the more vunerable it may be to pathogen infections causing defoliation.
However, such predictors would not be available for a spatial prediction scenario and one of the main goals of this study is to relate defoliation to variables that are available on a larger scale (e.g. remote sensing indices).

	[COMPARE ALL VARIABLES VS 7 VARIABLES]

@@

It is expected that models that were trained on the plot level only achieve a better performance than the global model.
The low performance on Laukiz 2 for the super model is most likely due to the difference of this plot to all others: The fitted model on Laukiz 1, Luiando and Oiartzun is not capable of reaching a good performance on the Laukiz 2 prediction data.
This is not surprising as Laukiz 2 shows substantially different plot characteristics compared to all others plots in terms of the distribution of the response variable \textit{defoliation} (\autoref{fig:defol_boxplots}) and the mean point density of trees (\autoref{fig:plot-characteristics}).

@@

There are some downsides using the internal variable importance approach of xgboost: Due to the contribution of three different parts to the overall importance score it is complicated to understand why a specific feature was selected.
Furthermore the importance calculation approach is only valid for this algorithm and cannot be compared to others.
Nevertheless, as we only relied on the variable importance for this specific algorithm, using the internal xgboost approach was sufficient for this work.

It is not surprising that vegetation indices are most important for the model as they are most sensitive to changes in vegetation health.
Even though we are not specifically looking at vegetation health but at the level of defoliation, these indices seem to help the model deciding whether a tree is defoliated or not.
The vegetation indeces could help here in two ways:
1) Trees that show a high level of defoliation do also reflect their bad health status through the remaining foliation.
2) Defoliated trees have more influence of bare ground information in their pixel values and will therefore be identified as defoliated by the model.

Even though no NRI made it among the most important variables in this study (stating that the first seven of this study are the most important ones), it is interesting that they occupy all ranks from 8 - 30 (\autoref{fig:var-imp}).
Our statement saying that the important indices of this study are the first seven can be seen critical as we only based the selection on a visual inspection of the variable importance results (\autoref{fig:var-imp}).
All indices starting from rank eight do simply have a very small score.
Nevertheless, one could also make the cut at rank five here.
However, the break between vegetation indeces and NRI at rank eight also had an influence in setting this subjective break there.
